---
title: "Final Q3"
author: "Leah  Castro Galvan"
date: "2026-05-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tigerstats)
library(tidyverse)
```

Final Q3.   This question is based on the GSS.csv data set.   The research questions is this:  "Is there a difference in the 2016 voter turnout based on region of the country?"

In the GSS.csv data set, Voted is the variable recording whether one voted in 2016 or not.
For purposes of this problem, compare regions SOUTH ATLANTIC, PACIFIC, AND MOUNTAIN

Your mission, should you decide to accept, is to analyze the GSS data for these three regions, and answer the research question. 

# Introduction
```{r}
GSSdata = read.csv('GSS1996.csv')
```
We begin by loading in out data GSS1996.csv so we can answer our research question: 
Is there a difference in the 2016 voter turnout based on region of the country?

# Methods

# Data Description

## Graphial Description 

## Numberical Description

# Inferential results

# Summary and Conclustions



